---
title: "Multi-Stage Area Probability Sample Design for the Detroit Transportation Security Survey"
subtitle: "SURV 745 - PracTools Project"
author:
  - "Team Neyman"
  - "Namit Shrivastava"
  - "Nicholas Reign Lugu"
date: "April 20, 2026"
format:
  pdf:
    documentclass: article
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
    fontsize: 11pt
    linestretch: 1
    number-sections: true
    toc: true
    toc-depth: 3
    fig-pos: "H"
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{fancyhdr}
        \pagestyle{fancy}
        \fancyhead[L]{SURV 745 - Team Neyman}
        \fancyhead[R]{Transportation Security Survey}
        \fancyfoot[C]{\thepage}
        \usepackage{booktabs}
        \usepackage{longtable}
        \usepackage{float}
execute:
  echo: true
  warning: false
  message: false
---

\newpage

# Introduction

This report presents the sample design for a face-to-face survey of residents of the
City of Detroit, Michigan. The survey aims to examine transportation security and
transportation use among Detroit residents. Key variables of interest include the
proportion of residents living in a household that does not own or lease a vehicle,
the proportion of residents whose main means of transportation is public
transportation, the proportion of residents who skip going somewhere because of a
problem with transportation, and the proportion of residents who asked for
transportation help from friends, family, or neighbors in the past month.

I designed a multi-stage area probability sample using census tracts as primary
sampling units (PSUs), block groups (BGs) as secondary sampling units (SSUs), and
persons as the ultimate sampling elements. The target sample size is 1,000
respondents distributed equally across four age groups (250 per group), with
domain-specific sampling rates that account for differential expected response rates.
The design achieves self-weighting within each age group and equal interviewer
workload across all sample PSUs.

The report is structured so that managers can follow the high-level logic of the
sampling and weighting plan in the main narrative sections, while technical personnel
can find detailed formulae, probabilities, and computational steps in the subsections
and appendices.

```{r setup}
#| label: setup
#| echo: true

# Load required libraries
library(tidyverse)    # Data manipulation and visualization
library(readxl)       # Reading Excel files
library(sampling)     # Survey sampling functions (systematic PPS)
library(knitr)        # Table formatting
library(kableExtra)   # Enhanced table formatting
library(sf)           # Spatial data handling
library(tigris)       # Census geographic shapefiles
library(ggplot2)      # Plotting
library(glue)         # String interpolation

# Set seed for reproducibility as specified in project instructions
set.seed(-77)

# Set tigris options
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
```

# Sample Design

## Goals of the Sample Design

The goals of the sample design are to:

1. Select a probability sample that yields approximately 1,000 responding adults
   (aged 18 and older) from the City of Detroit, with 250 respondents in each of
   four age groups (18--34, 35--49, 50--64, and 65+).

2. Achieve a self-weighting sample within each age group so that every person in a
   given age domain has the same overall probability of selection. This simplifies
   the weighting process and produces equal-weighted estimates within domains.

3. Obtain approximately equal interviewer workload in every sample PSU, so that
   field resources can be allocated uniformly across the 40 selected census tracts.

The following table summarizes the target sample sizes and expected response rates
by age group:

| Age Group | Target Respondents (n) | Expected Response Rate |
|-----------|----------------------:|----------------------:|
| 18--34    |                   250 |                  0.35 |
| 35--49    |                   250 |                  0.45 |
| 50--64    |                   250 |                  0.50 |
| 65+       |                   250 |                  0.65 |
| **Total** |             **1,000** |                       |

## Target Population

The target population consists of all civilian, non-institutionalized persons aged
18 years and older residing in the City of Detroit, Michigan. Detroit is located
within Wayne County, Michigan (FIPS state code 26, county code 163). According to
the 2020 Decennial Census, the city contains approximately 639,000 persons in
roughly 310,000 households.

## Area Sampling Frame: Units, Data Sources, and Construction

### Data Sources

I constructed the area sampling frame from two sources:

1. **Census Tract FIPS Code Detroit MI.csv**: A list of FIPS codes for all census
   tracts within the City of Detroit, obtained from the City of Detroit Open Data
   Portal (https://data.detroitmi.gov/).

2. **Detroit MI.xlsx**: Census tract-level and block-group-level population data from
   the 2020 Decennial Census Demographic and Housing Characteristics (DHC) Summary
   File. This file was prepared using the `tidycensus` R package to download
   Wayne County tract and block group level data from the Census API, then filtering
   to Detroit tracts. The data includes total household counts (H1_001N), total
   person counts, and person counts by sex and age from the P12 table.

### Frame Construction

```{r load-data}
#| label: load-data
#| echo: true

# Read the Census Tract FIPS codes for Detroit
TractFIPS <- read.csv("Census Tract FIPS Code Detroit MI.csv",
                       header = TRUE,
                       col.names = c("GEOID"),
                       colClasses = "character")

# Remove any blank or whitespace-only entries
TractFIPS <- TractFIPS %>%
  filter(trimws(GEOID) != "") %>%
  mutate(GEOID = trimws(GEOID))

# Get unique Detroit tract FIPS codes
detroit_tracts <- unique(TractFIPS$GEOID)
cat("Number of unique Detroit tract FIPS codes:", length(detroit_tracts), "\n")

# Read tract-level data from the Excel file
tract_data <- read_excel("Detroit MI.xlsx", sheet = "Tract") %>%
  mutate(Tract = as.character(Tract))

# Read block-group-level data
bg_data <- read_excel("Detroit MI.xlsx", sheet = "BlockGroup") %>%
  mutate(BlockGroup = as.character(BlockGroup),
         TractFIPS = substr(BlockGroup, 1, 11))

cat("Tracts in data:", nrow(tract_data), "\n")
cat("Block groups in data:", nrow(bg_data), "\n")
```

I created four age-domain variables by combining the detailed sex-by-age columns from
the P12 table. The domains correspond to the survey's four target age groups:

```{r age-groups}
#| label: age-groups
#| echo: true

# Function to create age group variables from the P12 sex-by-age columns
create_age_groups <- function(df) {
  df <- df %>%
    mutate(
      # Age 18-34: combine male and female counts for ages 18 through 34
      Age18to34 = Male18to19Yrs + Male20Yrs + Male21Yrs + Male22to24Yrs +
                  Male25to29Yrs + Male30to34Yrs +
                  Female18to19Yrs + Female20Yrs + Female21Yrs + Female22to24Yrs +
                  Female25to29Yrs + Female30to34Yrs,
      # Age 35-49
      Age35to49 = Male35to39Yrs + Male40to44Yrs + Male45to49Yrs +
                  Female35to39Yrs + Female40to44Yrs + Female45to49Yrs,
      # Age 50-64
      Age50to64 = Male50to54Yrs + Male55to59Yrs + Male60to61Yrs + Male62to64Yrs +
                  Female50to54Yrs + Female55to59Yrs + Female60to61Yrs + Female62to64Yrs,
      # Age 65+
      Age65plus = Male65to66Yrs + Male67to69Yrs + Male70to74Yrs + Male75to79Yrs +
                  Male80to84Yrs + MaleGE85Yrs +
                  Female65to66Yrs + Female67to69Yrs + Female70to74Yrs + Female75to79Yrs +
                  Female80to84Yrs + FemaleGE85Yrs,
      # Under 18 (for reference)
      AgeUnder18 = MaleLT5Yrs + Male5to9Yrs + Male10to14Yrs + Male15to17Yrs +
                   FemaleLT5Yrs + Female5to9Yrs + Female10to14Yrs + Female15to17Yrs,
      # Total adults 18+
      Adult18plus = Age18to34 + Age35to49 + Age50to64 + Age65plus
    )
  return(df)
}

# Apply to both tract and block group data
tract_data <- create_age_groups(tract_data)
bg_data <- create_age_groups(bg_data)
```

### Handling Tracts with Zero or Very Small Populations

```{r frame-cleaning}
#| label: frame-cleaning
#| echo: true

# Remove duplicate tract entries (tract 985500 appears 25 times as water body)
tract_dedup <- tract_data %>%
  distinct(Tract, .keep_all = TRUE)
cat("Tracts after removing duplicates:", nrow(tract_dedup), "\n")

# Identify tracts with zero adult population
zero_pop <- tract_dedup %>% filter(Adult18plus == 0)
cat("Tracts with zero adult population:", nrow(zero_pop), "\n")

# Identify tracts with very small adult populations (< 50)
small_pop <- tract_dedup %>% filter(Adult18plus > 0, Adult18plus < 50)
cat("Tracts with 1-49 adults:", nrow(small_pop), "\n")

# Remove zero and very small population tracts from the frame.
# These are typically water bodies, industrial areas, or institutional tracts
# (e.g., FIPS codes starting with 98xx) that have no residential population
# suitable for a household survey. Including them would create extreme weights
# and potential field operation problems.
tract_frame <- tract_dedup %>%
  filter(Adult18plus >= 50)
cat("Tracts in final frame:", nrow(tract_frame), "\n")

# Filter block groups to match the final tract frame
bg_frame <- bg_data %>%
  filter(TractFIPS %in% tract_frame$Tract)
cat("Block groups in final frame:", nrow(bg_frame), "\n")
```

After cleaning, the final frame contains 266 census tracts and 614 block groups.
I removed 10 tracts that had zero or near-zero adult population. These were
predominantly water-body tracts (FIPS codes in the 98xx range) and a few
special-purpose tracts with fewer than 50 adult residents. Retaining them would have
produced extreme sampling weights and created logistical difficulties during
fieldwork with no meaningful contribution to the survey estimates.

## Assigning Measures of Size

A key challenge in this design is simultaneously achieving self-weighting within each
age group and equal workload across PSUs, despite the four age groups having
different target sample sizes relative to their population shares and different
expected response rates. I addressed this by using a **composite measure of size**
(MOS).

### Deriving the Composite Measure of Size

For each age group $g$ ($g = 1, 2, 3, 4$ corresponding to 18--34, 35--49, 50--64,
65+), I first computed the overall sampling fraction needed to achieve the target
number of respondents:

$$f_g = \frac{n_g}{r_g \times N_g}$$

where $n_g = 250$ is the target number of respondents, $r_g$ is the expected response
rate, and $N_g$ is the total population in age group $g$ across the frame.

The composite MOS for any geographic unit (tract or block group) is then:

$$\text{MOS}_i = \sum_{g=1}^{4} f_g \times N_{g,i}$$

where $N_{g,i}$ is the population of age group $g$ in unit $i$. This composite MOS
reflects the expected workload (number of persons to be approached) in the unit,
weighted by the domain-specific sampling fractions.

```{r compute-mos}
#| label: compute-mos
#| echo: true

# Compute population totals for each age group across the frame
N1 <- sum(tract_frame$Age18to34)   # Total 18-34 population
N2 <- sum(tract_frame$Age35to49)   # Total 35-49 population
N3 <- sum(tract_frame$Age50to64)   # Total 50-64 population
N4 <- sum(tract_frame$Age65plus)   # Total 65+ population
N_adult <- sum(tract_frame$Adult18plus)  # Total adult 18+ population
N_total <- sum(tract_frame$TotPerson)    # Total population
N_hh <- sum(tract_frame$TotHH)          # Total households

cat("Frame population totals:\n")
cat("  Total persons:", N_total, "\n")
cat("  Total households:", N_hh, "\n")
cat("  Adults 18+:", N_adult, "\n")
cat("  Age 18-34:", N1, "\n")
cat("  Age 35-49:", N2, "\n")
cat("  Age 50-64:", N3, "\n")
cat("  Age 65+:  ", N4, "\n")

# Compute domain-specific sampling fractions
f1 <- 250 / (0.35 * N1)  # 18-34
f2 <- 250 / (0.45 * N2)  # 35-49
f3 <- 250 / (0.50 * N3)  # 50-64
f4 <- 250 / (0.65 * N4)  # 65+

cat("\nDomain-specific sampling fractions:\n")
cat("  f_1 (18-34):", round(f1, 6), "\n")
cat("  f_2 (35-49):", round(f2, 6), "\n")
cat("  f_3 (50-64):", round(f3, 6), "\n")
cat("  f_4 (65+):  ", round(f4, 6), "\n")

# Compute composite MOS for tracts
tract_frame <- tract_frame %>%
  mutate(CompMOS = f1 * Age18to34 + f2 * Age35to49 +
                   f3 * Age50to64 + f4 * Age65plus)

# Compute composite MOS for block groups
bg_frame <- bg_frame %>%
  mutate(CompMOS = f1 * Age18to34 + f2 * Age35to49 +
                   f3 * Age50to64 + f4 * Age65plus)

# Total composite MOS
MOS_total <- sum(tract_frame$CompMOS)
cat("\nTotal composite MOS:", round(MOS_total, 2), "\n")
cat("Expected workload per PSU:", round(MOS_total / 40, 2), "\n")
```

The composite MOS has an important property: when tracts are selected with
probabilities proportional to their composite MOS, and block groups within tracts are
also selected proportional to their composite MOS, the resulting design produces
exactly equal workload in every sample PSU and self-weighting within each age domain.

## Descriptive Statistics on the Frame

```{r frame-stats}
#| label: frame-stats
#| echo: true

# Summary statistics on the tract frame
frame_summary <- tract_frame %>%
  summarise(
    across(c(TotPerson, TotHH, Adult18plus, Age18to34, Age35to49,
             Age50to64, Age65plus, CompMOS),
           list(Mean = ~round(mean(.), 1),
                Median = ~round(median(.), 1),
                SD = ~round(sd(.), 1),
                Min = ~min(.),
                Max = ~max(.),
                Sum = ~sum(.)),
           .names = "{.col}_{.fn}")
  )

# Display as a formatted table
stats_df <- data.frame(
  Variable = c("Total Persons", "Total Households", "Adults 18+",
               "Age 18-34", "Age 35-49", "Age 50-64", "Age 65+",
               "Composite MOS"),
  Mean = c(frame_summary$TotPerson_Mean, frame_summary$TotHH_Mean,
           frame_summary$Adult18plus_Mean, frame_summary$Age18to34_Mean,
           frame_summary$Age35to49_Mean, frame_summary$Age50to64_Mean,
           frame_summary$Age65plus_Mean, frame_summary$CompMOS_Mean),
  Median = c(frame_summary$TotPerson_Median, frame_summary$TotHH_Median,
             frame_summary$Adult18plus_Median, frame_summary$Age18to34_Median,
             frame_summary$Age35to49_Median, frame_summary$Age50to64_Median,
             frame_summary$Age65plus_Median, frame_summary$CompMOS_Median),
  SD = c(frame_summary$TotPerson_SD, frame_summary$TotHH_SD,
         frame_summary$Adult18plus_SD, frame_summary$Age18to34_SD,
         frame_summary$Age35to49_SD, frame_summary$Age50to64_SD,
         frame_summary$Age65plus_SD, frame_summary$CompMOS_SD),
  Min = c(frame_summary$TotPerson_Min, frame_summary$TotHH_Min,
          frame_summary$Adult18plus_Min, frame_summary$Age18to34_Min,
          frame_summary$Age35to49_Min, frame_summary$Age50to64_Min,
          frame_summary$Age65plus_Min, frame_summary$CompMOS_Min),
  Max = c(frame_summary$TotPerson_Max, frame_summary$TotHH_Max,
          frame_summary$Adult18plus_Max, frame_summary$Age18to34_Max,
          frame_summary$Age35to49_Max, frame_summary$Age50to64_Max,
          frame_summary$Age65plus_Max, frame_summary$CompMOS_Max)
)

kable(stats_df, caption = "Descriptive Statistics for the Detroit Tract Frame (N = 266 tracts)",
      booktabs = TRUE, digits = 1) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

# Number of block groups per tract
bg_per_tract <- bg_frame %>%
  group_by(TractFIPS) %>%
  summarise(nBGs = n(), .groups = "drop")

cat("\nBlock groups per tract distribution:\n")
print(table(bg_per_tract$nBGs))

tract_frame <- tract_frame %>%
  left_join(bg_per_tract %>% rename(Tract = TractFIPS), by = "Tract")
```

The 266 tracts in the frame have an average of 2,403 total persons and 1,805 adults
aged 18+. The composite MOS ranges from a minimum of about 0.3 to a maximum of
19.5, reflecting substantial variation in tract size. The median tract has
approximately 2 block groups, though a few tracts have up to 6 block groups. Forty
tracts have exactly one block group, meaning the SSU selection probability for
those tracts will be 1.0 (the single BG is selected with certainty).

## Discussion: Equal Sampling Rate for All Persons

I now evaluate whether a design that uses the same sampling rate for all persons,
regardless of age group, could achieve the sample size targets.

```{r equal-rate}
#| label: equal-rate
#| echo: true

# If we use a single sampling rate f for all persons, the total expected
# respondents would be:
#   Total = f * (N1*r1 + N2*r2 + N3*r3 + N4*r4) = 1000
# So f = 1000 / sum(Ng * rg)

weighted_pop <- N1 * 0.35 + N2 * 0.45 + N3 * 0.50 + N4 * 0.65
f_equal <- 1000 / weighted_pop

cat("Equal sampling rate for all persons: f =", round(f_equal, 6), "\n\n")

# Expected respondents by age group under equal rate
equal_rate_results <- data.frame(
  AgeGroup = c("18-34", "35-49", "50-64", "65+"),
  Population = c(N1, N2, N3, N4),
  ResponseRate = c(0.35, 0.45, 0.50, 0.65),
  Target = rep(250, 4),
  Expected = round(c(f_equal * N1 * 0.35,
                      f_equal * N2 * 0.45,
                      f_equal * N3 * 0.50,
                      f_equal * N4 * 0.65), 1),
  Shortfall = round(c(f_equal * N1 * 0.35 - 250,
                        f_equal * N2 * 0.45 - 250,
                        f_equal * N3 * 0.50 - 250,
                        f_equal * N4 * 0.65 - 250), 1)
)

kable(equal_rate_results,
      caption = "Expected Respondents Under Equal Sampling Rate",
      booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

Under an equal sampling rate of approximately 0.00448, the expected number of
respondents would be 252.8 for the 18--34 group, 228.2 for the 35--49 group, 263.6
for the 50--64 group, and 255.3 for the 65+ group. The 35--49 age group would fall
about 22 respondents short of its target of 250, while the 50--64 group would be
oversampled by about 14.

The advantage of an equal-rate design is simplicity: every person has the same
probability of selection, which yields a self-weighting design overall (not just
within age domains). Weighting would be straightforward with a single base weight for
all respondents. However, the disadvantage is clear: it cannot simultaneously meet
all four age-group targets. The 35--49 group in particular would be underrepresented
because its population share relative to its response rate is the smallest among the
four groups.

For this reason, I used domain-specific sampling rates (i.e., the composite MOS
approach), which ensures that all four targets are met precisely while still
maintaining self-weighting within each age group.

# Sample Selection

## Method of Selection

I used **systematic sampling with probabilities proportional to size (systematic
PPS)** to select both the PSUs (tracts) and the SSUs (block groups). This method is
implemented in the `sampling` package in R via the `UPsystematic()` function.

The frame was sorted by census tract FIPS code prior to selection. Sorting by
geography provides an implicit geographic stratification effect, ensuring that the
sample is spread across the full extent of the city.

The selection proceeded in two stages:

- **Stage 1 (PSU selection):** 40 census tracts selected from the frame of 266
  tracts with probabilities proportional to their composite MOS.

- **Stage 2 (SSU selection):** Within each selected tract, 1 block group selected
  with probability proportional to its composite MOS within the tract.

```{r sample-selection}
#| label: sample-selection
#| echo: true

# Sort the frame by tract FIPS code before selection
tract_frame <- tract_frame %>% arrange(Tract)

# Compute inclusion probabilities for systematic PPS
# pi_i = m * MOS_i / sum(MOS) where m = 40
m <- 40
tract_frame <- tract_frame %>%
  mutate(pi_tract = m * CompMOS / sum(CompMOS))

# Check for any certainty selections (pi > 1)
cat("Max inclusion probability:", max(tract_frame$pi_tract), "\n")
cat("Min inclusion probability:", min(tract_frame$pi_tract), "\n")

# Cap at 1 if needed (not needed in our case since max < 1)
pik <- pmin(tract_frame$pi_tract, 1)

# Select 40 tracts using systematic PPS
# The UPsystematic function from the sampling package performs
# systematic sampling with unequal probabilities
set.seed(-77)
selected <- UPsystematic(pik)

tract_frame$selected <- selected
sample_tracts <- tract_frame %>% filter(selected == 1)
cat("Number of selected tracts:", nrow(sample_tracts), "\n")
```

## Stage 2: Block Group Selection Within Selected Tracts

```{r ssu-selection}
#| label: ssu-selection
#| echo: true

# For each selected tract, select 1 block group via PPS
set.seed(-77)  # Reset seed for SSU selection

sample_bgs <- data.frame()

for (i in 1:nrow(sample_tracts)) {
  tract_id <- sample_tracts$Tract[i]
  bgs_in_tract <- bg_frame %>%
    filter(TractFIPS == tract_id) %>%
    arrange(BlockGroup)

  n_bgs <- nrow(bgs_in_tract)

  if (n_bgs == 1) {
    # Only one BG in tract: selected with certainty
    sel_bg <- bgs_in_tract
    sel_bg$pi_bg_given_tract <- 1.0
  } else {
    # Compute within-tract inclusion probabilities for 1 BG
    tract_mos_total <- sum(bgs_in_tract$CompMOS)
    if (tract_mos_total > 0) {
      pik_bg <- bgs_in_tract$CompMOS / tract_mos_total
    } else {
      pik_bg <- rep(1 / n_bgs, n_bgs)
    }

    # Select 1 BG using systematic PPS
    sel_indicator <- UPsystematic(pik_bg)
    sel_bg <- bgs_in_tract[sel_indicator == 1, ]
    sel_bg$pi_bg_given_tract <- pik_bg[sel_indicator == 1]
  }

  # Add tract-level information
  sel_bg$TractID <- tract_id
  sel_bg$pi_tract <- sample_tracts$pi_tract[i]
  sel_bg$nBGs_in_tract <- n_bgs

  sample_bgs <- bind_rows(sample_bgs, sel_bg)
}

# Compute overall selection probability for each block group
sample_bgs <- sample_bgs %>%
  mutate(pi_overall = pi_tract * pi_bg_given_tract)

cat("Selected block groups:", nrow(sample_bgs), "\n")
```

## Selected Units and Their Characteristics

```{r sample-characteristics}
#| label: sample-characteristics
#| echo: true

# Display the selected PSUs with key characteristics
psu_table <- sample_tracts %>%
  select(Tract, NAME, TotPerson, Adult18plus, Age18to34, Age35to49,
         Age50to64, Age65plus, CompMOS, pi_tract, nBGs) %>%
  mutate(NAME = gsub("; Wayne County; Michigan", "", NAME),
         pi_tract = round(pi_tract, 4))

kable(psu_table %>% select(Tract, NAME, TotPerson, Adult18plus, CompMOS, pi_tract, nBGs),
      caption = "Selected PSUs (Census Tracts) and Their Characteristics",
      col.names = c("Tract FIPS", "Name", "Total Pop", "Adults 18+",
                     "Composite MOS", "Inclusion Prob", "No. BGs"),
      booktabs = TRUE, digits = c(0, 0, 0, 0, 2, 4, 0)) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                font_size = 9)
```

## Selection Probabilities at Each Stage

The selection probability structure for a person in age group $g$, residing in block
group $j$ within tract $i$, involves three stages:

**Stage 1 (Tract Selection):**
$$\pi_{1i} = m \times \frac{\text{MOS}_i}{\sum_{k=1}^{K}\text{MOS}_k} = 40 \times \frac{\text{MOS}_i}{`r round(MOS_total, 2)`}$$

**Stage 2 (Block Group Selection):**
$$\pi_{2j|i} = 1 \times \frac{\text{MOS}_{ij}}{\text{MOS}_i} = \frac{\text{MOS}_{ij}}{\sum_{l \in \text{tract } i}\text{MOS}_{il}}$$

**Stage 3 (Person Selection):**
$$f_{g,ij} = \frac{f_g}{\pi_{1i} \times \pi_{2j|i}}$$

The overall probability of selection for a person in age group $g$ is:
$$\pi_{g,ij} = \pi_{1i} \times \pi_{2j|i} \times f_{g,ij} = f_g$$

This confirms the self-weighting property: every person in age group $g$ has the
same overall selection probability $f_g$, regardless of which tract or block group
they reside in.

```{r person-rates}
#| label: person-rates
#| echo: true

# Compute person sampling rates within each selected BG for each domain
sample_bgs <- sample_bgs %>%
  mutate(
    f_person_18to34 = f1 / pi_overall,
    f_person_35to49 = f2 / pi_overall,
    f_person_50to64 = f3 / pi_overall,
    f_person_65plus = f4 / pi_overall,
    # Workload: total persons to approach in this BG
    workload = f_person_18to34 * Age18to34 +
               f_person_35to49 * Age35to49 +
               f_person_50to64 * Age50to64 +
               f_person_65plus * Age65plus
  )

cat("Workload per PSU (should be constant):\n")
cat("  Mean:", round(mean(sample_bgs$workload), 2), "\n")
cat("  SD:  ", round(sd(sample_bgs$workload), 4), "\n")
cat("  Min: ", round(min(sample_bgs$workload), 2), "\n")
cat("  Max: ", round(max(sample_bgs$workload), 2), "\n")

# Check for person sampling rates exceeding 1
for (col in c("f_person_18to34", "f_person_35to49",
              "f_person_50to64", "f_person_65plus")) {
  max_rate <- max(sample_bgs[[col]])
  cat(glue("  Max {col}: {round(max_rate, 4)}"), "\n")
}
```

## Description of How Persons Should Be Selected from Area Listings

Within each selected block group, interviewers will conduct a listing operation to
enumerate all dwelling units and the persons residing in them. Once the listing is
complete, persons should be selected using the domain-specific sampling rates
$f_{g,ij}$ computed above.

The procedure is as follows:

1. **Listing:** Enumerate all dwelling units in the selected block group and record
   the age (or age group) of each adult resident aged 18+.

2. **Classify persons by age domain:** Assign each listed adult to one of the four
   age groups (18--34, 35--49, 50--64, 65+).

3. **Apply domain-specific sampling rates:** Within each domain, select persons
   using systematic random sampling at the rate $f_{g,ij}$ specific to that block
   group and age domain. For example, if $f_{1,ij} = 0.05$ for the 18--34 group in
   a particular BG, then 1 in every 20 listed persons aged 18--34 would be selected.

4. **Within-household selection:** If multiple eligible adults from the same
   age domain reside in one household, select at most one per household using a
   random selection procedure (e.g., a Kish grid or next-birthday method).

# Anticipated Precision

I now estimate the anticipated precision of the key survey estimate: the proportion
of Detroit residents who report public transportation as their main means of
transportation.

```{r precision}
#| label: precision
#| echo: true

# Anticipated precision for the proportion using public transit
# Based on the Detroit Metro Area Communities Study, approximately 15-25%
# of Detroit residents rely on public transportation as their main mode.
# I use p = 0.15 as a conservative estimate for the anticipated proportion.

p_hat <- 0.15          # anticipated proportion
n_resp <- 1000         # total expected respondents
m_psu <- 40            # number of PSUs
b_bar <- n_resp / m_psu  # average cluster size per PSU

# Intraclass correlation: for geographic clustering of transit use
# within census tracts, a moderate rho of 0.02 is reasonable because
# transit use is influenced by local infrastructure, which varies across
# tracts but is somewhat homogeneous within a tract.
rho <- 0.02

# Design effect due to clustering
deff <- 1 + (b_bar - 1) * rho
cat("Design effect (DEFF):", round(deff, 2), "\n")

# Variance of estimated proportion under the cluster design
var_p <- deff * p_hat * (1 - p_hat) / n_resp
se_p <- sqrt(var_p)
cv_p <- se_p / p_hat

cat("Anticipated SE:", round(se_p, 4), "\n")
cat("Anticipated CV:", round(cv_p * 100, 1), "%\n")
cat("95% CI: (", round(p_hat - 1.96 * se_p, 4), ",",
    round(p_hat + 1.96 * se_p, 4), ")\n")

# Also compute for p = 0.25 (higher estimate scenario)
p2 <- 0.25
var_p2 <- deff * p2 * (1 - p2) / n_resp
se_p2 <- sqrt(var_p2)
cat("\nWith p = 0.25:\n")
cat("  SE:", round(se_p2, 4), ", CV:", round(se_p2/p2*100, 1), "%\n")
cat("  95% CI: (", round(p2 - 1.96*se_p2, 4), ",",
    round(p2 + 1.96*se_p2, 4), ")\n")

# Effective sample size
n_eff <- n_resp / deff
cat("\nEffective sample size:", round(n_eff, 0), "\n")
```

For an anticipated proportion of $p = 0.15$, the estimated standard error is 0.0137,
giving a coefficient of variation of 9.2% and a 95% confidence interval of
approximately (0.123, 0.177). The design effect is 1.48, which means the cluster
sample is roughly 48% less efficient than a simple random sample of the same size.
This corresponds to an effective sample size of about 676 persons.

If the true proportion is higher, say $p = 0.25$, the standard error increases to
0.0167 but the CV improves to 6.7%, which is well within acceptable precision for
most survey applications.

These calculations assume an intraclass correlation of $\rho = 0.02$, which is
reasonable for a behavioral outcome like transportation mode that shows moderate
geographic clustering.

# Description of Weighting Plan

The weighting plan has three components: base weights, nonresponse adjustments, and
post-stratification. I describe each in turn.

## Base Weights

The base weight for a responding person in age group $g$ from block group $j$ in
tract $i$ is the inverse of the overall selection probability:

$$w_{g,ij}^{base} = \frac{1}{\pi_{g,ij}} = \frac{1}{f_g}$$

Because the design is self-weighting within each age group, the base weight is
constant for all persons in the same age domain:

| Age Group | Sampling Fraction ($f_g$) | Base Weight ($1/f_g$) |
|-----------|:------------------------:|:--------------------:|
| 18--34    | 0.004427                 | 225.9                |
| 35--49    | 0.004905                 | 203.9                |
| 50--64    | 0.004246                 | 235.5                |
| 65+       | 0.004384                 | 228.1                |

## Nonresponse Adjustment

Nonresponse adjustments should be computed within weighting cells formed by the
cross-classification of sample PSU and age domain. Within each cell, the
nonresponse-adjusted weight is:

$$w_{g,ij}^{NR} = w_{g,ij}^{base} \times \frac{\sum_{k \in \text{selected}} w_{g,ij}^{base}}{\sum_{k \in \text{respondents}} w_{g,ij}^{base}}$$

If the weighting cells are too small (fewer than 20 respondents), cells should be
collapsed across adjacent PSUs or age groups before computing the adjustment factor.

## Post-Stratification

After nonresponse adjustment, post-stratification to known population totals from the
2020 Census (or the most recent American Community Survey estimates) should be
applied. The post-stratification dimensions should include at minimum age group
crossed with sex, which are available from the P12 table. If additional external
totals are available (e.g., race/ethnicity by age), they can be incorporated using
raking (iterative proportional fitting).

The final weight for a respondent is:

$$w_{g,ij}^{final} = w_{g,ij}^{NR} \times \frac{N_g^{post}}{{\hat{N}_g^{NR}}}$$

where $N_g^{post}$ is the known population control total for the post-stratification
cell and $\hat{N}_g^{NR}$ is the weighted estimate from the nonresponse-adjusted
weights.

# Variance Estimation

## Defining Analysis Strata and PSUs

For variance estimation, analysts should treat the 40 selected census tracts as
the PSUs and define analysis strata as needed. Because the design used systematic
PPS sampling without explicit stratification, a conservative approach is to
treat the entire sample as a single stratum. However, forming variance estimation
strata by pairing adjacent sample tracts (in the order they were selected from the
sorted frame) will provide a better variance estimate by capturing the implicit
stratification benefit.

I recommend the following approach:

1. **Form variance strata:** Create 20 variance strata by pairing the 40 sample
   tracts in the order they appear on the sorted frame. The first and second
   selected tracts form stratum 1, the third and fourth form stratum 2, and so on.

2. **Define PSUs as tracts:** Each tract is a PSU within its variance stratum.

3. **Use linearization (Taylor series) variance estimation:** Most survey software
   (R's `survey` package, SAS PROC SURVEYMEANS, Stata's `svy` prefix) can compute
   linearized variance estimates from the stratum and PSU identifiers along with
   the final weights.

```{r variance-setup}
#| label: variance-setup
#| echo: true

# Create variance estimation strata by pairing adjacent sample tracts
sample_bgs <- sample_bgs %>%
  arrange(TractID) %>%
  mutate(
    psu_order = row_number(),
    var_stratum = ceiling(psu_order / 2)
  )

cat("Variance strata created:", max(sample_bgs$var_stratum), "\n")
cat("PSUs per stratum: 2 (in each stratum)\n")

# Base weights for each age group
sample_bgs <- sample_bgs %>%
  mutate(
    base_wt_18to34 = 1 / f1,
    base_wt_35to49 = 1 / f2,
    base_wt_50to64 = 1 / f3,
    base_wt_65plus = 1 / f4
  )

# Display base weights
cat("\nBase weights by age group:\n")
cat("  18-34:", round(1/f1, 1), "\n")
cat("  35-49:", round(1/f2, 1), "\n")
cat("  50-64:", round(1/f3, 1), "\n")
cat("  65+:  ", round(1/f4, 1), "\n")
```

In R, the analysis would be set up as follows:

```r
library(survey)
design <- svydesign(
  ids = ~TractID,          # PSU identifier
  strata = ~var_stratum,   # Variance estimation stratum
  weights = ~final_weight, # Final analysis weight
  data = survey_data
)

# Estimate proportion using public transit
svymean(~public_transit, design)
```

# Maps

## Map of the City of Detroit

```{r detroit-map}
#| label: detroit-map
#| echo: true
#| fig-cap: "City of Detroit: Census Tracts (solid) and Block Groups (dotted)"
#| fig-width: 8
#| fig-height: 6

# Download tract and block group shapefiles from tigris
Detroit_tract <- tracts(state = "MI", county = "Wayne", year = 2020) %>%
  right_join(data.frame(GEOID = tract_frame$Tract), by = "GEOID")

Detroit_bg <- block_groups(state = "MI", county = "Wayne", year = 2020) %>%
  mutate(GEOIDt = substr(GEOID, 1, nchar(GEOID) - 1)) %>%
  right_join(data.frame(GEOID = tract_frame$Tract),
             by = c("GEOIDt" = "GEOID")) %>%
  select(-GEOIDt)

# Transform to WGS84 for mapping
Detroit_tract_trans <- st_transform(Detroit_tract, "+proj=longlat +datum=WGS84")
Detroit_bg_trans <- st_transform(Detroit_bg, "+proj=longlat +datum=WGS84")

# Plot all tracts and block groups
ggplot() +
  geom_sf(data = Detroit_bg_trans, fill = NA,
          linetype = "dotted", linewidth = 0.3) +
  geom_sf(data = Detroit_tract_trans, fill = NA,
          color = "black", linewidth = 0.6) +
  ggtitle("Detroit Tracts (Solid) and Block Groups (Dotted)") +
  theme_minimal() +
  theme(axis.text = element_text(size = 8))
```

## Map of Selected Tracts and Block Groups

```{r sample-map}
#| label: sample-map
#| echo: true
#| fig-cap: "Selected Sample Tracts (shaded blue) and Block Groups (shaded red)"
#| fig-width: 8
#| fig-height: 6

# Create indicator for selected tracts
Detroit_tract_trans <- Detroit_tract_trans %>%
  mutate(selected = ifelse(GEOID %in% sample_tracts$Tract, "Selected", "Not Selected"))

# Get selected BGs
selected_bg_ids <- sample_bgs$BlockGroup
Detroit_bg_trans <- Detroit_bg_trans %>%
  mutate(selected = ifelse(GEOID %in% selected_bg_ids, "Selected", "Not Selected"))

ggplot() +
  geom_sf(data = Detroit_tract_trans,
          aes(fill = selected), color = "black", linewidth = 0.4) +
  scale_fill_manual(values = c("Not Selected" = "white",
                                "Selected" = "steelblue"),
                    name = "Tracts") +
  geom_sf(data = Detroit_bg_trans %>% filter(selected == "Selected"),
          fill = "red", alpha = 0.5, color = "darkred", linewidth = 0.3) +
  ggtitle("Selected Sample: Tracts (Blue) and Block Groups (Red)") +
  theme_minimal() +
  theme(axis.text = element_text(size = 8),
        legend.position = "bottom")
```

\newpage

# Appendix A: Codebook for Frame and Sample Files

## Tract Frame File (FRAME1_NEYMAN)

| Variable | Description |
|----------|------------|
| Tract | 11-digit Census tract FIPS code (state + county + tract) |
| NAME | Census tract name |
| TotHH | Total households (from H1_001N) |
| TotPerson | Total persons |
| TotMale | Total male persons |
| TotFemale | Total female persons |
| MaleLT5Yrs ... MaleGE85Yrs | Male population by age group (from P12) |
| FemaleLT5Yrs ... FemaleGE85Yrs | Female population by age group (from P12) |
| Age18to34 | Persons aged 18-34 (both sexes combined) |
| Age35to49 | Persons aged 35-49 (both sexes combined) |
| Age50to64 | Persons aged 50-64 (both sexes combined) |
| Age65plus | Persons aged 65+ (both sexes combined) |
| AgeUnder18 | Persons under 18 (both sexes combined) |
| Adult18plus | Total adult population aged 18+ |
| CompMOS | Composite measure of size |
| nBGs | Number of block groups in the tract |

## Block Group Frame File (FRAME2_NEYMAN)

| Variable | Description |
|----------|------------|
| BlockGroup | 12-digit Census block group FIPS code |
| TractFIPS | 11-digit parent tract FIPS code |
| NAME | Block group name |
| TotHH ... FemaleGE85Yrs | Same demographic variables as tract file |
| Age18to34 ... Adult18plus | Same age-group variables as tract file |
| CompMOS | Composite measure of size |

## Sample File (SAMPLE_NEYMAN)

| Variable | Description |
|----------|------------|
| BlockGroup | 12-digit block group FIPS code of selected SSU |
| TractID | 11-digit parent tract FIPS code (selected PSU) |
| pi_tract | First-stage selection probability of the tract |
| pi_bg_given_tract | Second-stage conditional selection probability of the BG |
| pi_overall | Overall selection probability of the BG |
| f_person_18to34 | Within-BG person sampling rate for age 18-34 |
| f_person_35to49 | Within-BG person sampling rate for age 35-49 |
| f_person_50to64 | Within-BG person sampling rate for age 50-64 |
| f_person_65plus | Within-BG person sampling rate for age 65+ |
| base_wt_18to34 ... base_wt_65plus | Base weight for each age domain |
| workload | Expected number of persons to approach in this BG |
| var_stratum | Variance estimation stratum (1-20) |
| psu_order | PSU order number for variance strata assignment |
| All demographic variables | Same as block group frame |

\newpage

# Appendix B: Listing of Sample PSUs and SSUs

```{r appendix-listing}
#| label: appendix-listing
#| echo: false

# Create the appendix listing table
appendix_table <- sample_bgs %>%
  select(TractID, BlockGroup, TotPerson, Adult18plus,
         Age18to34, Age35to49, Age50to64, Age65plus,
         pi_tract, pi_bg_given_tract, pi_overall,
         f_person_18to34, f_person_35to49, f_person_50to64, f_person_65plus) %>%
  mutate(across(starts_with("pi_"), ~round(., 4)),
         across(starts_with("f_person_"), ~round(., 4)))

kable(appendix_table,
      caption = "Sample PSUs and SSUs with Selection Probabilities and Person Sampling Rates",
      booktabs = TRUE, longtable = TRUE,
      col.names = c("Tract", "Block Group", "Tot Pop", "Adults",
                     "18-34", "35-49", "50-64", "65+",
                     "pi(Tract)", "pi(BG|Tract)", "pi(Overall)",
                     "f(18-34)", "f(35-49)", "f(50-64)", "f(65+)")) %>%
  kable_styling(latex_options = c("hold_position", "scale_down", "repeat_header"),
                font_size = 7)
```

\newpage

# Appendix C: Output Frame and Sample Files

```{r write-outputs}
#| label: write-outputs
#| echo: true

# Write the tract frame file
frame1 <- tract_frame %>%
  select(Tract, NAME, TotHH, TotPerson, TotMale, TotFemale,
         Age18to34, Age35to49, Age50to64, Age65plus, AgeUnder18,
         Adult18plus, CompMOS, nBGs)
write.csv(frame1, "FRAME1_NEYMAN.csv", row.names = FALSE)
cat("FRAME1_NEYMAN.csv written:", nrow(frame1), "rows\n")

# Write the block group frame file
frame2 <- bg_frame %>%
  select(BlockGroup, TractFIPS, NAME, TotHH, TotPerson, TotMale, TotFemale,
         Age18to34, Age35to49, Age50to64, Age65plus, AgeUnder18,
         Adult18plus, CompMOS)
write.csv(frame2, "FRAME2_NEYMAN.csv", row.names = FALSE)
cat("FRAME2_NEYMAN.csv written:", nrow(frame2), "rows\n")

# Write the sample file
sample_out <- sample_bgs %>%
  select(BlockGroup, TractID, NAME, TotPerson, TotHH, Adult18plus,
         Age18to34, Age35to49, Age50to64, Age65plus,
         pi_tract, pi_bg_given_tract, pi_overall, nBGs_in_tract,
         f_person_18to34, f_person_35to49, f_person_50to64, f_person_65plus,
         base_wt_18to34, base_wt_35to49, base_wt_50to64, base_wt_65plus,
         workload, var_stratum, psu_order)
write.csv(sample_out, "SAMPLE_NEYMAN.csv", row.names = FALSE)
cat("SAMPLE_NEYMAN.csv written:", nrow(sample_out), "rows\n")
```

\newpage

# Appendix D: Disclosure of AI Use

This project used artificial intelligence tools to assist with code development,
report writing, and computational verification. Specifically:

**AI Model:** Claude (Anthropic), model claude-opus-4-6

**Purpose of AI Use:**

- Assisted in writing R code for frame construction, sample selection, weight
  computation, and map generation.
- Helped structure the sampling report and draft technical descriptions of the
  methodology.
- Verified numerical computations for composite measures of size, selection
  probabilities, and anticipated precision calculations.

**Prompts Used (Verbatim):**

The following prompts were used in conversations with the AI model:

1. "Read and analyze all the files in the folder for this project. I want proper R
   codes written with proper comments, inferences in a natural human like first
   person tone without the use of em dash. A sampling report (details of the report
   below); file name: REPORT_NEYMAN.pdf. Text files giving the units used for the
   area frames and relevant census counts and measures of size; file name:
   FRAME#_NEYMAN.pdf. Text file for the selected sample along with relevant census
   counts, measures of size, selection probabilities, and weights; file name:
   SAMPLE_NEYMAN.pdf."

2. "Properly generate a .qmd file with all the codes written properly and the report
   as per the instructions. Go through all the links provided in the Project.pdf and
   accurately generate the codes in the qmd file and in the report."

All AI-generated code and text were reviewed, verified, and edited by the team
members to ensure accuracy and appropriateness for the project requirements.
